### 2024-05-13 index.qmd
init

---
title: "Current Sheet Theory"
code-links:
#   - repo
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

## References

- @steinhauerMultifluidModelOnedimensional2008