### 2024-05-14 index.qmd
update

---
title: "Current Sheet Theory"
code-links:
  - repo
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

One-dimensional geometry in which all variables depend only on the z (normal) coordinate.

$$
d B_y / d z=\sum_\alpha(-n_\alpha u_{\alpha x}+\Gamma_\alpha B_x / B_z)
$$

$$
d B_x / d z=\sum_\alpha(n_\alpha u_{\alpha y}-\Gamma_\alpha B_y / B_z)
$$

$$
\Gamma_e=\sum_\alpha n_\alpha u_{\alpha z}=\sum_\alpha \Gamma_\alpha
$$

We are interested in solution that have $d B_y / d z \neq 0$

## References

- @steinhauerMultifluidModelOnedimensional2008