### 2024-10-04 index.qmd
update: SPARTHRB presentation

---
title: "Multifluid model of one-dimensional force-free current sheet equilibrium"
format:
  html: default
  pptx:
    slide-level: 2
    incremental: true
code-links:
  - text: Tana
    href: https://app.tana.inc?nodeid=OzzX0sVA4_6X
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
---

::: hide
\newcommand{\vect}[1]{\mathbf{#1}}
:::

# Introduction

Classification of discontinuities

The ratio $R_{V B}$ is expected to equal unity for rotational discontinuity, but is arbitrary for tangential discontinuities.

$$
R_{V B} = \frac{1}{\sqrt{A}} \frac{|Δ \vect{V}|}{|Δ \vect{V_A}|}
$$

where $Δ \vect{V_A} = Δ {\frac{\vect{B}}{\sqrt{\mu_0 \rho}}}$ and the anisotropy factor A is defined as $A = 1 - \frac{p_{\|}-p_{\perp}}{B^2/\mu_0}$.

## Examples

![Three examples of current sheets observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.](figures/fig_examples.svg)

---

## Force-free current sheets

Force-free current sheets are important for modelling low-$β$ plasmas in laboratory, space and astrophysical applications [@marshForcefreeMagneticFields1996]

However, observations of current sheet reveals discrepancies between

- Alfven velocity and plasma velocity change across current sheet
- Anisotropic MHD theory-predicted and directly measured ion anisotropies

## Theory - previous art

### Multi-fluid model

A multi-fluid model is developed to estimate the ratio of $R_{V B}$ [@neugebauerReexaminationRotationalTangential1984]. However no analytical solution is provided for the force-free current sheet.

$$
R_{V B}(\text { multistream })=\left(\frac{1}{A} \sum_i \frac{|Δ \vect{V_i}|^2}{|Δ \vect{V_{i,A}} |^2}\right)^{1 / 2}
$$

![Ion trajectories in model magnetic field.](figures/jgra18999-fig-0001.png)

---

### Steinhauer, McCarthy, and Whipple (2008)

@steinhauerMultifluidModelOnedimensional2008 built a multifluid model to study the steady state magnetotail current sheet.

For the symmetric three-fluid system with $B_y=0$. However, it is not force-free.

![Magnetic field, current density, ion density and out-of-plane flow for Bz = 0.1, βe = 0.5, βi = 1. Solid lines are fluid model results and dot symbols are from the Harris sheet solution.](figures/jgra18999-fig-0003.png)

---

### Vlasov equilibrium

A number of Vlasov-Maxwell distribution function have been found to describe force-free current sheet recently

$$
\begin{aligned}
& \frac{\mathrm{d}^2 A_x}{\mathrm{d} z^2}=-\mu_0 \frac{\partial P_{z z}}{\partial A_x} \\
& \frac{\mathrm{d}^2 A_y}{\mathrm{d} z^2}=-\mu_0 \frac{\partial P_{z z}}{\partial A_y}
\end{aligned}
$$

::: {.notes}
As has been shown before, the 1D VM equilibrium equations are equivalent to the motion of a pseudoparticle in a conservative pseudopotential, with the pseudopotential being proportional to one of the diagonal components of the plasma pressure tensor.
:::

- no normal field, $B_z = 0$

<!-- > Consider several deficiencies of the Harris model: (1) there is no normal field, Bz = 0, so that no particles cross the sheet from asymptotic sources; (2) the unrealistic assumption of uniform cross-tail drift velocity, uy; (3) vanishing particle density far from the sheet; (4) assumption of a simple Maxwellian velocity distribution; (5) further, one study showed that atypical current sheet structures with bifurcated current density occur 17% of the time, and even for typical center-peaked sheets, strongly peaked structures frequently appear in the center, which also differs from the Harris profile [Asano et al., 2005]. -->

# Plasma-Field Equations

## Multi-fluid collisionless plasma model in a nutshell

The governing equations for the multi-fluid collisionless plasma model can be expressed in conservation law form as follows:

$$
(\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) n_α = - n_α ∇ · \mathbf{u}_α
$$

$$
m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right)
$$

$$
\mathbf {\nabla } \times \mathbf {B} =\mu _{0}\mathbf {J}
$$ {#eq-Ampere}

$$
∇ · \mathbf{B} = 0
$$

where $α$ indicates the particle species...

---

## To much, let's simplify

- equilibrium solutions => $\frac{∂}{∂ t}=0$

- current sheet => magnetic field depends only on a single Cartesian coordinate, $z$

- force-free condition => $B_x^2 + B_y^2 = B_0^2 = \text{const}$

Additionally, we assume

- background electron is massless => electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$

- ion populations are gyrotropic => eliminating any non-diagonal terms in the pressure tensor

To further simplify, we employ a dimensionless system by normalizing the variables with their asymptotic values:

- magnetic field by $B_{\text{ref}} = B_z$

- density by $n_{\text{ref}} = n(∞)$

---

## After all the simplifications

Conservation of fluid mass integrates to a constant parameter $Γ_α = n_α u_{αz}$;

Gauss's law gives another constant parameter $B_{z} = \text{const}$;

The momentum equations for each ion species are then given by:

$$
\begin{aligned}
Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
\\
Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
\\
Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
\end{aligned}
$${#eq-momentum}

Ampere’s law connects the fields and the flow components and note that electron motion is along the field lines $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$ :

$$
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
$$ {#eq-Jx}

$$
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
$$ {#eq-Jy}

---

This is a system of $3N+2$ equations for $4N+2$ unknown dependent variables: $B_x$, $B_y$, and $N$ each of $n$, $p$ $u_x$, and $u_y$.

- $B_z, Γ_α$ are constant.

- Combine momentum equation (x) (@eq-momentum) and Ampere's law (y) (@eq-Jy) with condition that the constant of integration vanishes at the current sheet center yield

  $$
  B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
  $$

  - With asymptotic condition that the derivative goes zero, from momentum equation we have

    $$
    {B_z}^2 = \sum \Gamma_{\alpha}^2/n_{\alpha}(\infty)
    $${#eq-asym1}

In the following, we will consider a system with two ion species.

## Force-free current sheet

We are interested in solutions with $B_x^2 + B_y^2 = B_0^2 = const$ (force-free current sheet). This provides another equation for the system, now we have 3*2+2+1=9 equations for 10 unknowns (the system is close to be fully determined). From Ampere's law (@eq-Jx) and (@eq-Jy), we have:

$$
n (u_x B_y  - u_y B_x) = 0
$$

And $n_1 u_{1x} B_y = n_1 u_{1y} B_x + C_1$ and $n_2 u_{2x} B_y = n_2 u_{2y} B_x  - C_1$. Assuming $C_1 = 0$, from first two equations in (@eq-momentum), we have:

$$
u_{αx}^2 + u_{αy}^2 = const
$$

---

<!-- 
::: {#def-1}
Express the quantities of the second ion species relative to the first species

$$
u_{αx} := λ_{αx} u_{1x}, \quad u_{αy} := λ_{αy} u_{1y}, \quad u_{αz} := λ_{αz} u_{1z}, \quad n_α := λ_{αn} n_1
$$

$$
Λ_x := 1 + \sum λ_{αx} λ_{αn}, \quad Λ_y := 1 + \sum λ_{αy} λ_{αn}, \quad Λ_z := \sum 1 + λ_{αz} λ_{αn}
$$

Notes

- $Γ_e = (1 + λ_z λ_n) Λ_1 =  Λ_z  Γ_1$. 
- $Λ_z$ is constant while $Λ_x, Λ_y$ are not.
::: 
-->

::: {#def-2}
Force-free condition let us express the magnetic field and velocity in terms of the angle $θ$:

$$
B_x = B_0 \cos θ, \quad B_y = B_0 \sin θ
$$

$$
u_{1x} = u_{1} \cos θ_1, \quad u_{1y} = u_{1} \sin θ_1
$$
:::

---

The first momentum equation (@eq-momentum) after substituting the above expression becomes:

$$
- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
$$

$$
\Rightarrow 
θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
$${#eq-theta1}

- Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants determined by the system.
- **So given the profile of $n_α$, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.**
- The derivative of $θ$ goes to zero at infinity, gives us a relation between $u_α$ and $B_0$:

  $$
  u_α = \pm \frac{B_0 Γ_α}{B_z n_α(\infty)}
  $${#eq-u1}

The Ampere's law (@eq-Jy) become:

$$
- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
$$

$$
\Rightarrow 
θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
$${#eq-theta}

The above equation relates the rotation angle $θ$ to the plasma bulk velocity $n u \equiv \sum n_α u_{α}$. By equating the above two equations of $θ'$, we could get a equation relating the plasma bulk velocity to one specific species:

$$
n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
$$

## Solutions

The system $S$ could be fully determined by $λ_n, λ_z, B_0$ parameters, provided $n(z)$ profiles.

$$
S=S(z; n(z); λ_n, λ_z, B_0)
$$

Assuming density profile

$$
n_1(z) \to n_1(∞) +  \frac{κ Γ_1}{1+(z/L)^2}
$$

Solving the @eq-theta1-asym analytically, we have the rotation angle profile as

$$
\theta (z)\to \frac{\pi}{2} - \kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)
$$

---

## Beutiful moment - solutions

:::: {.columns}

::: column
In dimensionless form

$$
\begin{aligned}
& B_x(z) = B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& B_y(z) = B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& n(z) = n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
\\
& J_x(z) = \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& J_y(z) = \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$
:::

::: column
In physical units

$$
\begin{aligned}
& B_x(z) = B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& B_y(z) = B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& n(z) = n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
\\
& J_x(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& J_y(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$
:::
::::

---

For the simplest case $λ_n = 1, λ_z = - 1$, from @eq-asym1, we have $Γ_1 = - B_z / \sqrt{2}$.

We could also normalize the system length by $\delta = 1$. So now the system could be fully determined by $c, B_0$.

Given $c = 1/\sqrt{2}$, we have $B_y \to 0$ as $z \to \infty$. Profiles are plotted below for $B_0 = 2$.

![](figures/profiles.jpg)

![](figures/J_profiles_sym.svg)

The multi-fluid effect is clearly demonstrated by $Λ_y - 1$.

For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.

$$
\text{cond}(n(z), λ_{n,∞}, λ_{z,∞}) = 0
$$

Here we set $B_y(z=\infty) = 1/2 B_y(z=0)$ and $B_0 = 2 B_z$, and fix $λ_{z,∞} = -1$. By varying $λ_{n,∞}$, we find that the magnetic field profiles are exactly the same, while plasma velocity profiles vary. We normalize the plasma velocity by asympotic Alfvén velocity $v_{A,∞} = B_0 / \sqrt{n}$, and the profiles are plotted below @fig-uxnorm. It could be seen that for $λ_{n,∞} = 1$, we have zero bulk velocity change across the current sheet in the asymptotic limit.

---

## Symmetric case

![](figures/profiles_sym.svg){#fig-profilesEx1}

![](figures/J_profiles_sym.svg){#fig-jprofilesEx1}

(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $δ_1=1, c_1=1, B_0 = 2$. (b). Current density profiles for the same case.

## Asymmetric case

![](figures/profiles_n1Inf=0.6.svg)

![](figures/J_profiles_n1Inf=0.6.svg)

Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.

---

![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}

![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}

---

![Plasma velocity $U_x$ profiles normalized by $v_{A,x}$ for different $n_1(∞)$.](figures/UxNormBx.svg){#fig-uxNormBx}

![Plasma velocity $U_y$ profiles normalized by $v_{A,y}$ for different $n_1(∞)$.](figures/UyNormBy.svg){#fig-uyNormBy}

---

![](figures/vRatios.svg){#fig-vRatiosEx1}

![](figures/vRatiosEx2.svg){#fig-vRatiosEx2}

Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).