### 2024-10-04 article.qmd

---
title: On the Alfvenicity of solar wind discontinuities
authors:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
  - name: Sergei Kamaletdinov
    email: sergei2033@ucla.edu
    orcid: 0000-0002-2261-0331
    affiliation: University of California, Los Angeles
  - name: Xiaofei Shi
    email: sxf1698@ucla.edu
    orcid: 0000-0003-3367-5074
    affiliation: University of California, Los Angeles
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Marco Velli
    email: mvelli@ucla.edu
    orcid: 0000-0002-2381-3106
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
code-links:
  - text: Overleaf
    href: https://www.overleaf.com/project/668273f5d14067a22517615c
  - text: Roam Research
    href: https://roamresearch.com/#/offline/Research/page/5EiGYXsXE
  - text: Zotero
    href: zotero://select/library/collections/DZHNUFNV
abstract: "Solar wind plasma flow is characterized by intense, large-amplitude current sheets, also known as solar wind discontinuities. These regions are believed to play a crucial role in magnetic reconnection, solar wind ion heating, and energetic particle scattering. Although solar wind discontinuities share several properties with the classical MHD model of rotational discontinuities, they exhibit more complex characteristics that include many kinetic properties, going beyond this model. One of the most puzzling properties of these discontinuities is their Alvenicity, which describes the relationship between Alfvén and plasma velocity jumps across discontinuities. Although the high correlation of these jumps suggests the rotational nature of discontinuities, the significant differences in jump amplitudes indicate the necessity of kinetic mechanisms to describe the stress balance within discontinuities. Spacecraft observations show that plasma thermal anisotropy alone may not sufficiently explain this difference and achieve stress balance. In this study, we present a theoretical model in which this difference is explained by the kinetics of interpenetrating ion beams crossing the discontinuity. Our model successfully describes the observed properties of discontinuities within the inner heliosphere."
---

:::{.hide}
\newcommand{\vect}[1]{\mathbf{#1}}
:::

<!-- "This difference has previously been interpreted as: (1) a possible contribution of anisotropy that decreases ΔvA, or (2) a discontinuity non-stationarity due to residual magnetic energy. We propose an alternate interpretation: an ion nonadiabatic interaction with intense (thin) discontinuities that shapes the nongyrotropic ion distribution to include a nondiagonal term of the pressure tensor, with a cross-discontinuity gradient decreasing ΔvA." - @artemyevIonNongyrotropySolar2020 -->

# Introduction

- Observation features of solar wind discontinuities
- Importance of solar wind discontinuities
- The necessity of kinetic models to describe solar wind discontinuities
- Current theoretical models of solar wind discontinuities and their limitations
  - Harris current sheet @harrisPlasmaSheathSeparating1962

<!-- Observations of magnetic discontinuities exhibit Alfvenic properties, with the plasma velocity variation correlating well with magnetic field components variations. However, the amplitude of the plasma velocity jump is often significantly smaller than the Alfvén velocity jump, indicating a more complicated picture than the classical MHD model of rotational discontinuities. @fig-examples show observations of  magnetic discontinuities in different regions of the heliosphere, PSP within 0.2 AU, Wind upstream of the Earth's bow shock, and ARTEMIS in the Earth's magnetotail.  Previously, this discrepancy has been interpreted as a possible contribution of anisotropy that decreases the Alfvén velocity jump, or a discontinuity non-stationarity due to residual magnetic energy, or an ion nonadiabatic interaction with intense (thin) discontinuities that shapes the nongyrotropic ion distribution to include a nondiagonal term of the pressure tensor, with a cross-discontinuity gradient decreasing the Alfvén velocity jump @artemyevIonNongyrotropySolar2020.

These differences could not be simply explained by the aniostropy of the plasma thermal pressure alone [@shenComparingPlasmaAnisotropy2024] (assuming a Maxwellian distribution that fits both the parallel and perpendicular temperatures) and suggest that kinetic mechanisms are necessary to describe the stress balance within discontinuities. -->

::: {#fig-examples}
![](figures/fig_examples.svg)

Three examples of current sheets observed by Parker Solar Probe (PSP), ARTEMIS and Wind spacecraft with sub-Alfvenic flow velocity.
:::

<!-- Work on self-consistent kinetic models of the current sheet was pioneered by Sestero [1966] and by Schindler [1972, 1975, 1979] who presumed a much larger scale in the plane of the current sheet than in the normal (z) direction. Kan [1973] obtained exact solutions for the two dimensional (2-D) tail using a single Maxwellian distribution having uniform drift velocity and also a two component Maxwellian distribution having a sheared drift. His technique has been used by others to obtain 2-D solutions with an X-point (reviewed by Yoon and Lui [2005]). Eastwood [1972], Hill [1975], Cowley [1978a, 1978b], Cowley and Pellat [1979], Francfort and Pellat [1976], and others have discussed both the fluid and kinetic aspects of the current sheet, showing the importance of the balance between magnetic field pressure and the particle pressure component along the magnetic field, pk. -->

# Multifluid Model

We used a multi-fluid model for assumed 1-D quasineutral force-free current sheet equilibria. This approach, inspired by @steinhauerMultifluidModelOnedimensional2008, has the advantage of being analytically tractable while providing the necessary complexity to include the kinetic effects of multiple ion species. Kinetic approach, which is in principle more accurate, is hard to solve analytically and difficult to interpret the measurements. The multifluid model is a good compromise between the simplicity of the MHD model and the complexity of the kinetic model. One disadvantage of the multifluid model, however, is that there are more unknowns than equations. This means that the model is underdetermined and requires additional assumptions to be solved. This closure problem is usually solved by choosing equations of state. In this study, we first attempted to solve the system with a scalar pressure model, but then the system was overdetermined and no solution was found. We avoid this problem by assuming a specific profile for one of the variables, which is then used to solve the rest of the system.

In this model, we assume that the asymptotic boundary conditions are known from measurements, including the normal component of the magnetic field, the magnitude of the magnetic field, and the asymptotic properties of the incident fluid streams. It should be noted, however, that while the velocity space boundary between different groups may be clear in the uniform field region outside the current sheet, within the sheet itself, the boundary can become more complex. For instance, individual ions may jump from one group to another, and some may become trapped within the current sheet. These complexities, particularly ion behavior within the sheet, are beyond the scope of this paper and are not addressed in the current analysis.

<!-- An example is the simultaneous appearance of separate beams of ions in different directions with different fluxes. Parks et al. [2001] and Parks [2004] have pointed out that microphysical processes are key to understanding these new observations that, evidently, are unexplainable by conventional magnetohydrodynamic models. -->

# Plasma-Field Equations for Force-free current sheet

The governing equations for the multi-fluid collisionless plasma model can be expressed in conservation law form as follows:

$$
(\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) n_α = - n_α ∇ · \mathbf{u}_α
$$

$$
m_α n_α (\frac{∂}{∂ t} + \mathbf{u}_{α} · ∇) \mathbf{u}_α  = - ∇ · \mathbf{P}_α + q_α n_α \left( \mathbf{E} + \frac{\mathbf{u}_α \times \mathbf{B}}{c} \right)
$$

$$
\mathbf {\nabla } \times \mathbf {B} =\mu _{0}\mathbf {J}
$$ {#eq-Ampere}

$$
∇ · \mathbf{B} = 0
$$

where $α$ indicates the particle species, $n_α$ the number density, $m_α$ the mass, $q_α$ the charge, $\mathbf{u}_α$ the bulk velocity, $\mathbf{P}_α$ the pressure tensor, $\mathbf{E}$ the electric field a $\mathbf{B}$ the magnetic induction (magnetic field).

In this study, we focus on the equilibrium solutions of the set of equations, which implies that the partial derivatives with respect to time are set to zero ($\frac{∂}{∂ t}=0$). Since our primary interest lies in the force-free current sheet, we assume that the magnetic field depends only on a single Cartesian coordinate, $z$, leading to the force-free condition $B_x^2 + B_y^2 = B_0^2 = \text{const}$. Additionally, we assume that the background electron is massless, meaning the electrons follow the magnetic field lines $\mathbf{u}_e \times \mathbf{B} = 0$. For further simplification, we also assume the ion populations are gyrotropic, eliminating any non-diagonal terms in the pressure tensor.

Under these assumptions, the conservation of fluid mass, expressed as $d(n_α u_{αz})/dz = 0$, integrates to a constant parameter, $Γ_α = n_α u_{αz}$. The electron motion is aligned with the magnetic field lines, resulting in $\vect{u_e} = Γ_e \vect{B} / (n_e B_z)$. The momentum equations for each ion species are then given by:

$$
\begin{aligned}
Γ_α \frac{d u_{α x}}{d z} &= n_α u_{α y} B_z - Γ_α B_y
\\
Γ_α \frac{d u_{α y}}{d z} &= Γ_α B_x-n_α u_{α x} B_z
\\
Γ_α \frac{d u_{α z}}{d z} &= -\frac{1}{2} \frac{d p_α}{d z}-n_α \frac{d \phi}{d z}+n_α u_{α x} B_y-n_α u_{α y} B_x
\end{aligned}
$$ {#eq-momentum}

Ampere’s law connects the fields and the flow components

$$
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
$$ {#eq-Jx}

$$
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
$$ {#eq-Jy}

$$
0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
$$ {#eq-Jz}

And the Gauss's law for magnetism gives another constant parameter $B_{z} = \text{const}$.

This is a system of $3N+2$ equations for $4N+2$ unknown dependent variables: $B_x$, $B_y$, and $N$ each of $n$, $p$, $u_x$, and $u_y$. Since we are interested in force-free solutions with $B_x^2 + B_y^2 = B_0^2 = \text{const}$, this condition provides an additional equation for the system.

A scalar pressure model, which provides another N equations connecting $p$ and $n$, was attempted as per @steinhauerMultifluidModelOnedimensional2008, but no solution was found (the system was overdetermined). Therefore, the system is considered with free parameters. For a system with two ion species, there are 9 equations for 10 unknowns. The system can be solved by setting one of the parameters to have a specific profile and solving the rest of the system.

The force-free condition let us express the magnetic field in terms of the rotation angle $θ$ as $B_x = B_0 \cos θ$ and $B_y = B_0 \sin θ$. From Ampere's law (@eq-Jx) and (@eq-Jy), we have:

$$
n (u_x B_y  - u_y B_x) = 0
$$

And we could express the density and bulk velocity in terms of every ion species: $n_α u_{αx} B_y = n_α u_{αy} B_x + C_α$. This above equation can be rewritten as $\sum_α C_α =0$. The simplest case is to set $C_α=0$ for all species, which is the case we will consider in this study.

$$
n_α (u_{αx} B_y - n_α u_{αy} B_x) = 0
$$ {#eq-alpha}

Combining the above equation with the first two equations in (@eq-momentum), we have:

$$
u_{αx}^2 + u_{αy}^2 = const
$$

This would let us express the bulk velocity in terms of the rotation angle $θ_α$ as $u_{αx} = u_{α} \cos θ_α$ and $u_{αy} = u_{α} \sin θ_α$. And substituting the above expression into the our assumption (@eq-alpha), we have:

$$
\tan θ = \tan θ_α
$$

__#TODO#__ Here we consider the case where $θ_α = θ$ for all species, but the same procedure can be applied to the case where $θ_α = θ + \pi$ and it would give the same results.

The first momentum equation (@eq-momentum) after substituting the above expression becomes:

$$
- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
$$

$$
\Rightarrow 
θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
$${#eq-theta1}

Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system. So given the profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.

The Ampere's law (@eq-Jy) after substitution becomes:

$$
- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
$$

$$
\Rightarrow 
θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
$${#eq-theta}

The above equation relates the rotation angle $θ$ to the plasma bulk velocity $n u \equiv \sum n_α u_{α}$. By equating the above two equations of $θ'$, we could get a equation relating the plasma bulk velocity to one specific species:

$$
n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
$$

In the asymptotic region, as all variables approach constant values, the derivatives must vanish. For @eq-theta1, this means $θ_α(∞)' = 0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$. This relates the velocity of each species to the asymptotic density. Rewriting the @eq-theta1 in terms of asymptotic values, we have:

$$
θ_α' = \frac{B_z}{Γ_α} (n_α - n_α(∞))
$${#eq-theta1-asym}

Conveniently, the center of the current sheet is chosen as the origin $z = 0$, which corresponds to the lower boundary of @eq-theta1. As a result, the boundary condition for the rotation angle is given by  $\theta_α(0) = \pi/2$.

Combine momentum equation (x) (@eq-momentum) and Ampere's law (y) (@eq-Jy) with condition that the constant of integration vanishes at the current sheet center yield

$$
B_x B_z = \sum n_α u_{α,x} u_{α,z} = \sum Γ_α u_{α,x}
$$

Evaluate the above equation in the asymptotic limit, we have

$$
{B_z}^2 = \sum Γ_{α}^2/n_{α}(∞)
$${#eq-asym1}

To simplify the analysis, it is useful to employ a dimensionless system by normalizing the variables with their asymptotic values: in this study, the magnetic field is normalized by $B_{\text{ref}} = B_z$ , and the density is normalized by $n_{\text{ref}} = n(∞)$. Other reference values derived from these two are displayed in the table below.

| Variable | Reference Value |
| --- | --- |
| Frequency (ion plasma frequency) | $\omega_{pi} = \sqrt{\frac{n_{\text{ref}} e^2}{m_i \varepsilon_0}}$ |
| Length (ion inertial length) | $L_{\text{ref}} = c / \omega_{pi}$ |
| Velocity (Alfvén velocity) | $V_{\text{ref}} = B_{\text{ref}} / \sqrt{\mu_0 m_i n_{\text{ref}}}$ |

# Results

The multifluid model developed above is used here to study a specific example and the effects of model parameters: assume a density profile for one species taking the form of a Lorentzian function with a background density.

$$
n_1(z) \to n_1(∞) +  \frac{κ Γ_1}{1+(z/L)^2}
$$

Solving the @eq-theta1-asym analytically, we have the rotation angle profile as

$$
\theta (z)\to \frac{\pi}{2} - \kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)
$$

And therefore

$$
\begin{aligned}
B_x(z) &= B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
B_y(z) &= B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
n(z) &= n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
\\
J_x(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
J_y(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$

For the simplest situation where we have two group of ions of the same density but opposite bulk velocity, i.e., $n_1 = n_2, u_1 = -u_2$, the system could be determined given $κ$, $L$, and $B_0$. The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($L=1, κ=1, B_0 = 2$) is plotted below @fig-profiles. The specific profiles are chosen to have zero electron current across the current sheet and zero $B_y$ in the asymptotic limit, corresponding to a 180° rotation of the magnetic field across the current sheet. However, in general, we would expect the electron current to be non-zero.

::: {#fig-profiles layout-ncol="2" }
![](figures/profiles_sym.svg){#fig-profilesEx1}

![](figures/J_profiles_sym.svg){#fig-jprofilesEx1}

(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $L=1, κ=1, B_0 = 2$. (b). Current density profiles for the same case.
:::

## Alfvenicity

For the same asymptotic magnetic field, it is interesting to see how the plasma profiles change with different system parameters.

Here we set $B_y(z=\infty) = 1/2 B_0, B_0 = 2 B_z$, and let two ion species have the opposite bulk velocity but this time with different densities. By varying $n_1(∞)$, we find that the magnetic field profiles are exactly the same, while plasma density and velocity profiles vary.

::: {#fig-profilesEx2 layout-ncol="2" }
![](figures/profiles_n1Inf=0.6.svg)

![](figures/J_profiles_n1Inf=0.6.svg)

Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
:::

We normalize the plasma velocity by asympotic Alfvén velocity $v_{A}(∞) = B_0 / \sqrt{\mu_0 m_i n}$, and the profiles are plotted in @fig-profilesEx2 for the case of $n_1(∞) = 0.6$. 
@fig-uxnorm It could be seen that for $n_1(∞) = 0.5$, we have zero bulk velocity change across the current sheet in the asymptotic limit. And the change of bulk velocity across the current sheet increases with $n_1(∞)$ decreasing to zero. The normalized plasma velocity $U_y$ in the aympotic limit decreases with $n_1(∞)$ decreasing to zero @fig-uynorm.

::: {#fig-profilesC layout-nrow="1" }
![Plasma velocity $U_x$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UxNormB0.svg){#fig-uxnorm}

![Plasma velocity $U_y$ profiles normalized by aysmpotic Alfvén velocity for different $n_1(∞)$.](figures/UyNormB0.svg){#fig-uynorm}
:::

Normalized by Alfvén velocity $v_{A,x}$ and $v_{A,y}$, the plasma velocity profiles are plotted below @fig-uxNormBx and @fig-uyNormBy. The normalized plasma velocity profiles are exactly the same for $U_x/v_{A,x}$ and $U_y/v_{A,y}$ for same $n_1(∞)$. When $n_1(∞)$ decreases to zero, the normalized plasma velocity would approach one in the asymptotic limit, indicating the Alvenicity of the current sheet.

<!-- ::: {.column-page} -->
::: {#fig-uProfilesNorm layout-nrow="1" }
![Plasma velocity $U_x$ profiles normalized by $v_{A,x}$ for different $n_1(∞)$.](figures/UxNormBx.svg){#fig-uxNormBx}

![Plasma velocity $U_y$ profiles normalized by $v_{A,y}$ for different $n_1(∞)$.](figures/UyNormBy.svg){#fig-uyNormBy}
:::
<!-- ::: -->

The asymptotic velocity for the each species, bulk velocity and normalized one by Alfvén velocity are plotted below @fig-vRatios and @fig-vRatiosEx2 for the case of $u_{1,z} = - u_{2,z}$ (two groups of opposite velocity in z direction) and $u_{1,z} = - 2 u_{2,z}$ respectively. For the case of $u_{1,z} = - u_{2,z}$ as shown in @fig-vRatiosEx1, the normalized velocity reaches zero when $n_1(∞) = 0.5$. The bulk velocity of each species is the same in the asymptotic limit and does not depend on $n_1(∞)$. However, this does not generally hold when $u_{1,z} \neq - u_{2,z}$ as shown in @fig-vRatiosEx2 for the case of $u_{1,z} = - 2 u_{2,z}$. The normalized velocity reaches zero when $n_1(∞) = 1/3$, and the bulk velocity of each species differs in the asymptotic limit and depends on $n_1(∞)$.

<!-- ::: {.column-page} -->
::: {#fig-vRatios layout-nrow="1" }
![](figures/vRatios.svg){#fig-vRatiosEx1}

![](figures/vRatiosEx2.svg){#fig-vRatiosEx2}

Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).
:::
<!-- ::: -->

## Current Sheet Thickness

TODO: define and compare current sheet thickness

TODO: compare with Harris current sheet solution with same thickness

## Density Peaking

# Summary

In non-dimensionalized form
$$
\begin{aligned}
B_x(z) &= B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
B_y(z) &= B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
n(z) &= n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
\\
J_x(z) &= \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
J_y(z) &= \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$

::: {#fig-vRatioStats}
![](https://raw.githubusercontent.com/Beforerr/psp_conjunction/5e09973ebedbdaef5cc89cb3a12be3fb77f44570/figures/enc7/vl_ratio.png)

Statistics of the asymptotic velocity ratio from PSP, Wind, and ARTEMIS spacecraft observations during PSP encounter 7 period from 2021-01-14 to 2021-01-21.
:::