### 2024-11-18 presentations/AGU24.qmd
feat: add poster

---
title: Multifluid model of one-dimensional force-free current sheet equilibrium
subtitle: AGU24 poster
author:
  - name: Zijin Zhang
    affil: 1
  - name: Anton Artemyev
    affil: 1
  - name: Marco Velli
    affil: 1
  - name: Vassilis Angelopoulos
    affil: 1
affiliation:
  - num: 1
    address: Department of Earth, Planetary, and Space Sciences, University of California, Los Angeles
output: "posterdown::posterdown_betterland"
---

## Introduction & Motivation

'Discontinuities' are discontinuous spatial changes in plasma
parameters/characteristics and magnetic fields.

## Method

## Results

$$
\begin{aligned}
& B_x(z) = B_0 \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& B_y(z) = B_0 \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& n(z) = n_{\infty } + \frac{\kappa}{V_A} \frac{∑_α Γ_α}{1+(z/L)^2}
\\
& J_x(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \sin \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
& J_y(z) = \frac{\kappa B_0}{\mu_0 d_i}\frac{1}{1+(z/L)^2} \cos \left(\frac{\kappa L}{d_i} \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$

::: {#fig-profiles layout-ncol="2" }
![](../figures/profiles_sym.svg){#fig-profilesEx1}

![](../figures/J_profiles_sym.svg){#fig-jprofilesEx1}

(a). Magnetic field, ion density, and ion bulk velocity for $n_1 = n_2, u_1 = -u_2$ and $L=1, κ=1, B_0 = 2$. (b). Current density profiles for the same case.
:::

::: {#fig-profilesEx2 layout-ncol="2" }
![](../figures/profiles_n1Inf=0.6.svg)

![](../figures/J_profiles_n1Inf=0.6.svg)

Same as @fig-profiles, but for two ion species with same bulk velocity but different densities $n_1(∞) = 0.6$ and $B_y(z=\infty) = 1/2 B_0$.
:::

::: {#fig-vRatios layout-nrow="1" }
![](../figures/vRatios.svg){#fig-vRatiosEx1}

![](../figures/vRatiosEx2.svg){#fig-vRatiosEx2}

Asymptotic velocity for each species, bulk velocity and normalized one by Alfvén velocity for the case of $u_{1,z} = - u_{2,z}$ (a) and $u_{1,z} = - 2 u_{2,z}$ (b).
:::